---
title: "Analysis"
output: html_document
date: "2023-06-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(skimr)
library(janitor)
library(here)
library(chron)
```

```{r}
daily_activity <- read.csv("G:/Otros ordenadores/Mi Portátil/Gaspar Facultad/Analisis de datos/Caso practico/R/Boards/daily_activity_csv.csv")
```

```{r}
sleep_day <- read.csv("G:/Otros ordenadores/Mi Portátil/Gaspar Facultad/Analisis de datos/Caso practico/R/Boards/sleep_day_csv.csv")
```

# Analysis

## Frequency of use

The first step was to create a frequency table of device usage. This table considers the days on which each ID recorded information.

```{r}
sleep_day_frequency <- read.csv("G:/Otros ordenadores/Mi Portátil/Gaspar Facultad/Analisis de datos/Caso practico/R/Boards/daily_activity_frequency_csv.csv")
```

```{r}
daily_activity_frequency <- data.frame(daily_activity %>%
                                         group_by(Id) %>%
                                         summarize(min_date = min(ActivityDate),
                                                   max_date = max(ActivityDate)),
                                       table(daily_activity$Id))
```

```{r}
daily_activity_frequency <- daily_activity_frequency %>%
  select(-Var1)
```

```{r}
daily_activity_frequency
```

## Frequency of use during sleep

We proceeded to create an identical table, incorporating the information we have about users' sleep.

```{r}
sleep_day_frequency <- read.csv("G:/Otros ordenadores/Mi Portátil/Gaspar Facultad/Analisis de datos/Caso practico/R/Boards/sleep_day_frequency_csv.csv")
```

```{r}
str(sleep_day_frequency)
```
### Avg. days of use

```{r}
mean(daily_activity_frequency[ ,4])
```

```{r}
mean(daily_activity_frequency[ ,4])/31
```

On average, users use the device on 92% of the days during the study period, which is equivalent to 28 out of 31 days.

### Avg. days of device usage during sleep

```{r}
mean(sleep_day_frequency[ ,4])
```

```{r}
mean(sleep_day_frequency[ ,4])/31
```

On average, the device is used during sleep for only 17 out of 31 days, which is approximately 55%.

## Usage of the device by minutes of activity

Create a data frame that contains only the columns of ID, the different activity times according to their intensity, and a last column with the total sum of minutes of device usage.

```{r}
activity_minutes <- read.csv("G:/Otros ordenadores/Mi Portátil/Gaspar Facultad/Analisis de datos/Caso practico/R/Boards/activity_minutes_csv.csv")

```

```{r}
head(activity_minutes)
```

Averages of device usage in minutes are calculated based on the intensity of the activity.

```{r}
activity_minutes_avg <- data.frame(VeryActiveMinutes = round(mean(activity_minutes$VeryActiveMinutes)),
           ModeratelyActiveMinutes = round(mean(activity_minutes$ModeratelyActiveMinutes)),
           LightlyActiveMinutes = round(mean(activity_minutes$LightlyActiveMinutes)),
           SedentaryMinutes = round(mean(activity_minutes$SedentaryMinutes))
           )
```

```{r}
activity_minutes_avg
```

```{r}
activity_minutes_avg_percent <- read.csv("G:/Otros ordenadores/Mi Portátil/Gaspar Facultad/Analisis de datos/Caso practico/R/Boards/activity_minutes_avg_percent_csv.csv")
```

```{r}
activity_minutes_avg_percent
```

## Coincidencias

### Activity Data and Sleep Data

To analyze the activity data in conjunction with the sleep data, a new data frame is created. It contains the activity data only for users who also recorded sleep activity.

Merge the columns of interest from each data frame.

```{r}
coincidences <- merge(sleep_day ,
      daily_activity %>% 
        select(-VeryActiveDistance,
               -ModeratelyActiveDistance,
               -LightActiveDistance,
               -SedentaryActiveDistance),
      by.x = c("Id", "ActivityDate"), by.y = c("Id","ActivityDate"))
```

An additional column is added to the data frame with the total sum of activity minutes.

```{r}
coincidences <- read.csv("G:/Otros ordenadores/Mi Portátil/Gaspar Facultad/Analisis de datos/Caso practico/R/Boards/coincidence_csv.csv")
```

```{r}
coincidences <- coincidences %>% 
  mutate(TotalActiveMinutes = 
           VeryActiveMinutes + 
           ModeratelyActiveMinutes +
           LightlyActiveMinutes +
           SedentaryMinutes)
```

```{r}
head(coincidences)
```

Finally, a correlation coefficient table is created to understand the relationship that exists between the variables.

```{r}
correlation <- read.csv("G:/Otros ordenadores/Mi Portátil/Gaspar Facultad/Analisis de datos/Caso practico/R/Boards/correlation_csv.csv")
```

```{r}
correlation <- mutate_all(correlation, as.numeric)
```

```{r}
correlation <- cor(correlation)
```

```{r}
(correlation)
```
